Viscous Erosion in Porous Media

### 2018-02-02 docs/methods/erosion_methods.tex
round of edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion is a fluid-mechanical process that is responsible for a variety
of phenomena in geophysical fluids and other processes.  For example,
wind erosion causes sand dunes to be  longer on the windward
side~\cite{han1969}, and erosion is responsible for branching structures
of streams and rivers~\cite{coh-dev-sey-yi-szy-rot2015}.  Erosion is
also important in the biophysical community. Plaque erosion in arteries
is key to understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}.  Another application,
which motivates the current work, is erosion caused by groundwater flow.
Here, a slow moving fluid erodes channels in the ground, and this
affects the fluid transport.  In all these examples, there is a feedback
process between the flow and the eroding bodies---the shape of the
bodies affect the flow, while the flow changes the shape by fluidizing
each body.  This fluid-mechanical feedback system has been simulated in
idealized cases such as a single eroding two-dimensional body in a
uniform flow~\cite{moo-ris-chi-zha-she2013} or a single mobile eroding
three-dimensional body~\cite{mit-spa2016}. However, to better understand
the effect of erosion on groundwater flow, multiply-connected domains
must be considered.  Here we simulate two-dimensional erosion, in
multiply-connected domain such as the porous geometry of many
groundwater flows (Figure~\ref{fig:50bodies}).

@@

To create high-fidelity and computationally efficient numerical methods,
two fast and accurate methods are employed.  First, the fluid equations
and the erosion rate (shear stress) are efficiently solved with spectral
accuracy by using a fast multipole accelerated boundary integral
equation method.  Given the rate of erosion, the shape is integrated in
time with second-order accuracy.  All spatial derivatives are be
computed in Fourier space, resulting in spectral accuracy in space.
Then, to further visualize the effect of erosion, we compute the
vorticity of the flow (Figure~\ref{fig:50bodies}) to visualize regions
of high and low erosion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress}
\label{sec:shearStressLP}
The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential, the Stokeslets, and the
rotlets.  We compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right),
  \end{aligned}
\end{equation}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is the point chosen inside body
$\gamma_\ell$, and $\rho = \|\rr_\ell\|$.  Now that the deformation
tensor of each component of the completed double-layer potential is
computed, the shear stress at $\xx_0 \in \bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

@@

To handle boundary evolution, we use the {\thL} framework, which offers
certain advantages in the ability to stabilize fluid-structure
interaction problems \cite{hou-low-she1994}. Throughout this section, we
use the following conventions: the bodies are parametrized in the CCW
direction and the normal vector points into the bodies (or out of the
fluid).

For the sake of numerical stability, we modify the erosion law in
equation~\eqref{eqn:erosionModel} to include a term that depends on the
local curvature $\kappa$. The idea is to prevent regions of very high
curvature. The modified normal velocity is 
\begin{align}
  \label{eqn:Vn}
  \Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1 \right)
\end{align}
Here, $\eps$ is a small parameter, $L$ is the total arc length of the
body, and $\mean{\cdot}$ indicates a boundary mean. The second term on
the right acts to smooth bodies with high-frequency oscillations by
introducing a {\em curvature}-dependent component in the interface
velocity. Since a closed body has a mean curvature of $\mean{\kappa} =
2\pi / L$, the term in parentheses has mean zero. Thus, this term is
area-preserving, so that the only material loss is due to the shear
stress itself. In addition, we have scaled this term on the
$\mean{\atau}$, the average absolute stress, so that it is always
proportional to the true source of erosion $\atau$, even as bodies
vanish. \todo[inline]{reword}

We now introduce the tangent angle, $\theta$, defined through the relation
\begin{align*}
\left( \pderiv{x}{s}, \pderiv{y}{s} \right) = \left(\cos \theta, \sin
\theta \right),
\end{align*}
where $s$ is arc length. We also introduce the normalized arc length $\alpha = 2 \pi s / L$. In terms of these new variables, the curvature can be expressed as
\begin{align*}
\kappa = \frac{2 \pi}{L} \pderiv{\theta}{\alpha}.
\end{align*}
Substitution into~\eqref{eqn:Vn} gives
\begin{align}
\label{eqn:Vn2}
\Vn = \atau +  \eps \mean{\atau}   \left(\pderiv{\theta}{\alpha} - 1 \right)
%\Vn = \atau + \frac{2 \pi \eps}{L \log \left(2\pi/L \right)}  \left(\pderiv{\theta}{\alpha} - 1 \right)
\end{align}

The main idea is to introduce an artificial tangential velocity that
keeps collocation points equally spaced with respect to arc length;
i.e.~the grid of $\alpha$ should always remain equally spaced. The
tangential velocity that does the trick satisfies~\cite{hou-low-she1994}
\begin{align*}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}.
\end{align*}
The interface evolution is then given by
\begin{align*}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{align*}
This evolution equation can be recast in the variables $\theta$ and $L$
as
\begin{align*}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align*}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} + \thalpha \Vs \right)
\end{align}
Note that stiffest term in~\eqref{eqn:thetaevo1} involving the
second-order derivative is {\em quasi-linear}. This is the key to the
{\thL} method, and allows us to use very stable, implicit methods for
this stiff term that are described in Section~\ref{sec:timeStepping}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{s:method}
There are three main numerical methods that need to be developed.  In
Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE}
is solved.  Once the density function is found,
Section~\ref{sec:shearStress} describes how the shear
stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}
and~\eqref{eqn:thetaevo1} are described.

So that we can form stable simulations with long time horizons, we use
numerical methods that achieve spectral accuracy in space, second-order
accuracy in time, and maintain an equispaced discretization of the
bodies.  The boundaries are represented in Fourier space, and all
derivatives are computed with spectral accuracy using Fourier
differentiation.

@@

To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, we let $N_\iin$ be the number of
points on each of the $M$ bodies, and $N_\out$ be the number of points
on the outer boundary.  Then, $N = MN_\iin + N_\out$ is the total number
of discretization points of $\bd\Omega$.  The double-layer potential
in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The equations for the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{40pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_i \in \gamma_\ell} \eeta(\xx_i) 
      \Delta s_i, \\ 
    \mu_\ell &= \sum_{\xx_i \in \gamma_\ell}
      {(\xx_i - \cc_\ell)}^\perp \cdot \eeta(\xx_i) \Delta s_i,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  Since all the integrands are
smooth and periodic, by using the trapezoid rule, we are guaranteed that
the numerical solutions converges to the true solution with spectral
accuracy.

The linear system~\eqref{eqn:completed_BIE_discretized} is solved
iteratively with the generalized minimal residual method
(GMRES)~\cite{saa-sch1986}.  Since the BIE is second-kind, only a
mesh-independent number of GMRES iterations are
required~\cite{cam-ips-kel-mey-xue1996}.  Therefore, the asymptotic cost
is determined by the cost of performing a matrix-vector multiplication,
and the bulk of the cost is evaluating the $N$-term sum
in~\eqref{eqn:completed_BIE_discretized}.  Naively, evaluating this sum
at all points $\xx_i$, $i=1,\ldots,N$ requires $\bigO(N^2)$ operations.
To reduce the computational cost, we use the Fast Multipole Method
(FMM)~\cite{gre-rok1987, gre-gre-may1992} to reduce the complexity from
$\bigO(N^2)$ to $\bigO(N)$.  Since the number of required GMRES
iterations is mesh-independent, the overall complexity for
solving~\eqref{eqn:completed_BIE_discretized} is $\bigO(N)$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion is a fluid-mechanical process that is responsible for a variety
of phenomena in geophysical fluids and other processes.  For example,
wind erosion causes sand dunes to be  longer on the windward
side~\cite{han1969}, and erosion is responsible for branching structures
of streams and rivers~\cite{coh-dev-sey-yi-szy-rot2015}.  Erosion is
also important in the biophysical community. Plaque erosion in arteries
is key to understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}.  Another application,
which motivates the current work, is erosion caused by groundwater flow.
Here, a slow moving fluid erodes channels in the ground, and this
affects the fluid transport.  In all these examples, there is a feedback
process between the flow and the eroding bodies---the shape of the
bodies affect the flow, while the flow changes the shape by fluidizing
each body.  This fluid-mechanical feedback system has been simulated in
idealized cases such as a single eroding two-dimensional body in a
uniform flow~\cite{moo-ris-chi-zha-she2013} or a single mobile eroding
three-dimensional body~\cite{mit-spa2016}. However, to better understand
the effect of erosion on groundwater flow, multiply-connected domains
must be considered.  Here we simulate two-dimensional erosion, in
multiply-connected domain such as the porous geometry of many
groundwater flows (Fig.~\ref{fig:50bodies}).

@@

To create high-fidelity and computationally efficient numerical methods,
two fast and accurate methods are employed.  First, the fluid equations
and the erosion rate (shear stress) are efficiently solved with spectral
accuracy by using a fast multipole accelerated boundary integral
equation method.  Given the rate of erosion, the shape is integrated in
time with second-order accuracy.  All spatial derivatives are be
computed in Fourier space, resulting in spectral accuracy in space.
Then, to further visualize the effect of erosion, we compute the
vorticity of the flow (Fig.~\ref{fig:50bodies}) to visualize regions
of high and low erosion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress}
\label{sec:shearStressLP}
To compute the shear stress, there are two options.  First, instead of
solving the BIE~\eqref{eqn:completed_BIE} for an arbitrary density
function, a similar BIE can be solved for the
traction~\cite{mit-spa2016}.  Alternatively, the traction corresponding
to the double-layer potential of the velocity~\eqref{eqn:velocityDLP}
can be computed using appropriate layer potentials.  While the first
method bypasses a step by not forming a density function, this method
does not allow us to compute other quantities such as the vorticity.
Therefore, we use the second method.

The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Fig.~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential, the Stokeslets, and the
rotlets.  We compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$
is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given by
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right).
  \end{aligned}
\end{equation}
Now that the deformation tensor of each component of the completed
double-layer potential is computed, the shear stress at $\xx_0 \in
\bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

@@

To handle boundary evolution, we use the {\thL} framework, which offers
certain advantages in the ability to stabilize fluid-structure
interaction problems \cite{hou-low-she1994}. Throughout this section, we
use the following conventions: the bodies are parametrized in the
counterclockwise direction and the normal vector points into the bodies
(Fig.~\ref{fig:schematic}).

For the sake of numerical stability, we modify the erosion law in
equation~\eqref{eqn:erosionModel} to include a term that depends on the
local curvature $\kappa$. The idea is to prevent regions of very high
curvature. The modified normal velocity is 
\begin{align}
  \label{eqn:Vn}
  \Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1
  \right).
\end{align}
Here, $\eps$ is a small parameter, $L$ is the total arc length of the
body, and $\mean{\cdot}$ indicates a boundary mean. The second term on
the right acts to smooth bodies with high-frequency oscillations by
introducing a {\em curvature}-dependent component in the interface
velocity. Since a closed body has a mean curvature of $\mean{\kappa} =
2\pi / L$, the term in parentheses has mean zero. Thus, this term is
area-preserving, so that the only material loss is due to the shear
stress itself. In addition, we have scaled this term on the
$\mean{\atau}$, the average absolute stress, so that it is always
proportional to the true source of erosion $\atau$, even as bodies
vanish. \todo[inline]{reword}

We now introduce the tangent angle, $\theta$, defined through the relation
\begin{align*}
\left( \pderiv{x}{s}, \pderiv{y}{s} \right) = \left(\cos \theta, \sin
\theta \right),
\end{align*}
where $s$ is arc length. We also introduce the normalized arc length $\alpha = 2 \pi s / L$. In terms of these new variables, the curvature can be expressed as
\begin{align*}
\kappa = \frac{2 \pi}{L} \pderiv{\theta}{\alpha}.
\end{align*}
Substitution into~\eqref{eqn:Vn} gives
\begin{align}
\label{eqn:Vn2}
\Vn = \atau +  \eps \mean{\atau}   \left(\pderiv{\theta}{\alpha} - 1
\right).
%\Vn = \atau + \frac{2 \pi \eps}{L \log \left(2\pi/L \right)}  \left(\pderiv{\theta}{\alpha} - 1 \right)
\end{align}

The main idea is to introduce an artificial tangential velocity that
keeps collocation points equally spaced with respect to arc length;
i.e.~the grid of $\alpha$ should always remain equally spaced. The
tangential velocity that does the trick satisfies~\cite{hou-low-she1994}
\begin{align*}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn}.
\end{align*}
The interface evolution is then given by
\begin{align*}
\dot{\xx}(t) = \Vn \nn + \Vs \ss.
\end{align*}
This evolution equation can be recast in the variables $\theta$ and $L$
as
\begin{align*}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align*}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
\thalpha \Vs \right).
\end{align}
Note that stiffest term in~\eqref{eqn:thetaevo1} involving the
second-order derivative is {\em quasi-linear}. This is the key to the
{\thL} method, and allows us to use very stable, implicit methods for
this stiff term that are described in Section~\ref{sec:timeStepping}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{s:method}
There are three main numerical methods that need to be developed.  In
Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE}
is solved.  Once the density function is found,
Section~\ref{sec:shearStress} describes how the shear
stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}
and~\eqref{eqn:thetaevo1} are described. So that we can form stable
simulations with long time horizons, we use numerical methods that
achieve spectral accuracy in space, second-order accuracy in time, and
maintain an equispaced discretization of the bodies.  The boundaries are
represented in Fourier space, giving us a set of discretization points
$\xx_i$ of each body, and all derivatives and integrals (the arclength,
curvature, and mean values) are computed with spectral accuracy using
the Fourier representation.

@@

To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, let $N_\iin$ be the number of
points on each of the $M$ bodies, and $N_\out$ be the number of points
on the outer boundary.  Then, $N = MN_\iin + N_\out$ is the total number
of discretization points of $\bd\Omega$.  The double-layer potential
in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The equations for the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{53pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_j \in \gamma_\ell} \eeta(\xx_j) 
      \Delta s_j, \\ 
    \mu_\ell &= \sum_{\xx_j \in \gamma_\ell}
      {(\xx_j - \cc_\ell)}^\perp \cdot \eeta(\xx_j) \Delta s_j,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  Since all the integrands are
smooth and periodic, by using the trapezoid rule, we are guaranteed that
the numerical solutions converges to the true solution with spectral
accuracy.

The linear system~\eqref{eqn:completed_BIE_discretized} is solved
iteratively with the generalized minimal residual method
(GMRES)~\cite{saa-sch1986}.  Since the BIE is second-kind, only a
mesh-independent number of GMRES iterations are
required~\cite{cam-ips-kel-mey-xue1996}.  Therefore, the asymptotic cost
is determined by the cost of performing a matrix-vector multiplication,
and the bulk of the cost is evaluating the $N$-term sum
in~\eqref{eqn:completed_BIE_discretized}.  Naively, evaluating this sum
at all points requires $\bigO(N^2)$ operations.  To reduce the
computational cost, we use the Fast Multipole Method
(FMM)~\cite{gre-rok1987, gre-gre-may1992} to reduce the complexity from
$\bigO(N^2)$ to $\bigO(N)$.  Since the number of required GMRES
iterations is mesh-independent, the overall complexity for
solving~\eqref{eqn:completed_BIE_discretized} is $\bigO(N)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shear Stress}
\label{sec:shearStress}
We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  For the layer potential term, we would like to use the
trapezoid rule, but the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  Therefore, the deformation tensor of
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_\ell$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_\ell}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_\ell \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets are given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is computed with Fourier
differentiation.  Then, the shear stress is computed by multiplying the
deformation tensor by the normal vector, and taking the inner product
with the tangent vector.
